Concurrent Programming with CXL

### 2020-05-18 book/paper.tex
Latest

To see that $I_p(i)$ implies mutual exclusion, suppose not.  Then
when both process 0 and process 1 are in the critical section, the
following must hold:
$(\lnot C(0) \lor \mathtt{process}(1)@gate) \land
 (\lnot C(1) \lor \mathtt{process}(0)@gate)$.
We know that both flags are set.
We also know that neither process 1 nor process 2 is at label \texttt{@gate}
(because they are both at label \texttt{@cs}),
so this simplifies to $(\lnot C(0)) \land (\lnot C(1))$.
So we conclude that $\mathtt{turn} = 0 \land \mathtt{turn} = 1$, a
logical impossibility.  Thus $I_p(i)$ satisfies ME1.

@@

Finally we have to show that $I_p(i)$ satisfies ME3. Without loss of
generality, suppose $i=0$ (a benefit from the fact that the algorithm is
symmetric for both processes).  We have to show that if we are in a state
in which $I_p(0)$ holds, then any step will result in a state in which
$I_p(0)$ continues to hold.
Since $I_p(0)$ holds, process 0 is at label \texttt{@cs}.  If process 0
were to take a step, then in the next state process 0 would be no longer
at that label and $I_p(0)$ would hold trivially over the next state.
Therefore we only need to consider a step by process 1.

From $I_p(0)$ we know that one of the following cases must hold before
process 1 takes a step:
\begin{itemize}
\item \texttt{flags[1] = False};
\item \texttt{turn = 0};
\item process 1 is at label \texttt{@gate}.
\end{itemize}

Let us consider each of these cases.
In the first case, if process 1 takes a step, there are two possibilities:
either $flags[1]$ will still be \texttt{False} (in which case the first case
continues to hold), or $flags[1]$ will be \texttt{True}
and process 1 will be at label \texttt{@gate} (in which case the third case
will hold).
We know that process 1 never sets \texttt{turn} to 1, so
if the second case holds before the step, it will also hold after the step.
Finally, if process 1 is at label \texttt{@gate} before the step, then after
the step \texttt{turn} will equal 0, and therefore the second case will hold
after the step.

@@

@@

>>>>

To see that $I_p(i)$ implies mutual exclusion, suppose not.  Then
when both process~0 and process~1 are in the critical section, the
following must hold:
$(\lnot C(0) \lor \mathtt{process}(1)@gate) \land
 (\lnot C(1) \lor \mathtt{process}(0)@gate)$.
We know that both flags are set.
We also know that neither process~0 nor process~1 is at label \texttt{@gate}
(because they are both at label \texttt{@cs}),
so this simplifies to $(\lnot C(0)) \land (\lnot C(1))$.
So we conclude that $\mathtt{turn} = 0 \land \mathtt{turn} = 1$, a
logical impossibility.  Thus $I_p(i)$ satisfies ME1.

@@

Finally we have to show that $I_p(i)$ satisfies ME3. Without loss of
generality, suppose $i=0$ (a benefit from the fact that the algorithm is
symmetric for both processes).  We have to show that if we are in a state
in which $I_p(0)$ holds, then any step will result in a state in which
$I_p(0)$ continues to hold.
Since $I_p(0)$ holds, process~0 is at label \texttt{@cs}.  If process~0
were to take a step, then in the next state process~0 would be no longer
at that label and $I_p(0)$ would hold trivially over the next state.
Therefore we only need to consider a step by process~1.

From $I_p(0)$ we know that one of the following cases must hold before
process~1 takes a step:
\begin{itemize}
\item \texttt{flags[1] = False};
\item \texttt{turn = 0};
\item process~1 is at label \texttt{@gate}.
\end{itemize}

Let us consider each of these cases.
In the first case, if process~1 takes a step, there are two possibilities:
either $flags[1]$ will still be \texttt{False} (in which case the first case
continues to hold), or $flags[1]$ will be \texttt{True}
and process~1 will be at label \texttt{@gate} (in which case the third case
will hold).
We know that process~1 never sets \texttt{turn} to 1, so
if the second case holds before the step, it will also hold after the step.
Finally, if process~1 is at label \texttt{@gate} before the step, then after
the step \texttt{turn} will equal 0, and therefore the second case will hold
after the step.

@@

\begin{quote}
Bonus question!
\begin{itemize}
\item Can you generalize Peterson's algorithm to more than two processes?
\end{itemize}
\end{quote}

@@

Figure~\ref{fig:uplock} showed a faulty attempt at solving mutual exclusion.
